\ttitle

### 2016-04-16 response_letter.tex

% Chapter 01 - Introduction

\chapter*{An investigation of the clinical utility of preoperative cardiopulmonary exercise testing in patients undergoing major pancreatic surgery.}

@@

\lhead{\emph{Response to examiners' comments}} % This is for the header on each page - perhaps a shortened title

\vfill

\hrule
\vfill
{\LARGE \textbf{Response to examiners' comments}}
\vfill
\hrule

\vfill

@@

	Point taken.	
	The following paragraph has been included under Section 1.4.4 in Chapter 1 and discusses some of the key randomised trials of operative strategies aimed at reducing postoperative complications after pancreaticoduodenectomy.

	Several studies have attempted to evaluate various surgical strategies in improving perioperative outcomes after pancreaticoduodenectomy. 
	A prospective, multi-center, randomised controlled trial of 170 patients undergoing pancreaticoduodenectomy did not demonstrate any difference between a Whipple's classical pancreaticoduodenectomy  and pylorus preserving pancreaticoduodenectomy \parencite{tran_pylorus_2004}. 
	A Cochrane review in 2008 reported on 7 randomised controlled trials comparing classical pancreaticoduodenectomy with pylorus-preserving pancreaticoduodenectomy  and found no difference in morbidity, mortality or survival \parencite{diener_pancreaticoduodenectomy_2008}.

	A prospective randomised trial comparing pancreaticogastrostomy versus pancreaticojejunostomy in patients undergoing pancreaticoduodenectomy reported that there was no difference in the incidence of overall complications or pancreatic fistula. 
	However, patients who had a pancreaticogastrostomy had a significantly lower rate of biliary fistula, postoperative collections, delayed gastric emptying as well as lower incidence of multiple surgical complications \parencite{bassi_reconstruction_2005}. 
	A meta-analysis of eleven studies (one randomised trial, 2 non-randomised prospective trials and eight observational studies) appeared to show improved outcomes in patients undergoing pancreaticogastrostomy rather than a pancreaticojejunostomy \parencite{mckay_meta-analysis_2006}. 
	However, an earlier trial by Yeo and co-workers failed to demonstrate any difference in outcomes between either technique \parencite{yeo_prospective_1995}.

	In spite of a lack of any clear evidence of efficacy, octreotide continues to be used in many centres including ours to reduce the incidence of postoperative pancreatic fistula. 
	While some early randomised trials reported the benefits of octreotide in decreasing the risk of postoperative pancreatic fistula \parencite{montorsi_efficacy_1995,nakatsuka_octreotide_2000}, other more recent trials have not reproduced these results \parencite{lowy_prospective_1997, yeo_does_2000, kollmar_prophylactic_2008}.

@@

	Two randomised trials comparing early versus delayed drain removal in patients with low risk of pancreatic fistula after pancreaticoduodenectomy reported that early drain removal was associated with lower incidence of pancreatic fistula, abdominal complications and pulmonary complications \parencite{kawai_early_2006,bassi_early_2010}.

	However, these pancreas-specific risk factors are not modifiable before surgery, have been studied extensively before and are not the subject of this thesis.

	The following has been added to Section 1.3 in Chapter 1 and adds to the already detailed discussion of some of the key randomised trials in adjuvant therapy after pancreaticoduodenectomy.

	Early randomised trials reported improved survival in patients with inoperable pancreatic cancer \parencite{mallinson_chemotherapy_1980} as well as in patients who had undergone potentially curative surgery \parencite{bakkevold_adjuvant_1993}.

	Oettle and co-workers reported on the results of a European multi-center, randomised controlled phase 3 trial comparing the results of surgery verus surgery and adjuvant chemotherapy with 6 cycles of gemcitabine in patients with pancreatic ductal adenocarcinoma \parencite{oettle_adjuvant_2007}. 
	The median disease free survival in patient who received gemcitabine chemotherapy was 13.4 months while it was only 6.9 months in the surgery-only group. 
	However, there was no difference in overall survival.

	Increasingly, neo-adjuvant therapies are being used in patients with initially unresectable or 'borderline-resectable' cancers in an attempt to improve resectability. 
	In a meta-analysis of 111 studies, Gillen and co-workers reported that approximately one-third of initially unresectable patients can be expected to have resectable tumours after neo-adjuvant treatment \parencite{gillen_preoperative/neoadjuvant_2010}. 
	The authors recommended that such patients should be actively included in neo-adjuvant protocols and re-staged for resection after treatment.

### 2016-06-05 Chapters/Introduction.tex
Spellcheck - Introduction

\subsection{Postoperative care}
All patients were routinely admitted to the Surgical High Dependency Unit unless intra-operative events necessitated admission to the Intensive Care Unit. 
A standardised regimen of intravenous fluids, naso-jejunal feeding, mobilisation and physiotherapy was implemented in all patients. 
Standard physiological parameters including haemodynamic measurements, renal function and arterial blood gases were used to monitor adequate end organ perfusion. 
All patients received proton pump inhibitors and Octreotide. 
Patients were discharged to the general surgical ward as early as possible. 
Amylase content in the drain fluid was measured on the third post-operative day to aid in the diagnosis of post-operative pancreatic fistula. 
Surgical drains and naso-gastric/naso-jeujunal tubes were removed at the surgeon's discretion. 
Patients were discharged when medically fit.

@@

Oettle and co-workers reported on the results of a European multi-center, randomised controlled phase 3 trial comparing the results of surgery verus surgery and adjuvant chemotherapy with 6 cycles of gemcitabine in patients with pancreatic ductal adenocarcinoma \parencite{oettle_adjuvant_2007}. The median disease free survival in patient who received gemcitabine chemotherapy was 13.4 months while it was only 6.9 months in the surgery-only group. However, there was no difference in overall survival.

@@

Several studies have attempted to evaluate various surgical strategies in improving perioperative outcomes after pancreaticoduodenectomy. 
A prospective, multi-center, randomised controlled trial of 170 patients undergoing pancreaticoduodenectomy did not demonstrate any difference between a Whipple's classical pancreaticoduodenectomy  and pylorus preserving pancreaticoduodenectomy \parencite{tran_pylorus_2004}. 
A Cochrane review in 2008 reported on 7 randomised controlled trials comparing classical pancreaticoduodenectomy with pylorus-preserving pancreaticoduodenectomy  and found no difference in morbidity, mortality or survival \parencite{diener_pancreaticoduodenectomy_2008}.

@@

In spite of clear evidence of efficacy, octreotide continues to be used in many centres including ours to reduce the incidence of postoperative pancreatic fistula. 
While some early randomised trials reported the benefits of octreotide in decreasing the risk of postoperative pancreatic fistula \parencite{montorsi_efficacy_1995,nakatsuka_octreotide_2000}, other more recent trials have not reproduced these results \parencite{lowy_prospective_1997, yeo_does_2000, kollmar_prophylactic_2008}.

@@

$\dot{V}_{O_2}$ or oxygen uptake is measured breath-by-breath using digital analysis of the inspired and expired gases. 
This is then averaged, usually over time, to smooth-out any signficant breath-by-breath variation. $\dot{V}_{O_2}$ increases with increasing work load and is influenced by several factors that have a role in the transport and utilisation of oxygen. 
These may be broadly classified as cardiac, pulmonary, circulatory and tissue factors. 
Some of the factors are encompassed in the following formula for $\dot{V}_{O_2}$.
\begin{equation} \label{eq:oxygen_consumption}
	\dot{V}_{O_2} = CaO_2 \times Cardiac\ Output
\end{equation}
where $CaO_2$ is $O_2$ content per ml of blood and is defined by,
\begin{equation} \label{eq:CaO2}
	CaO_2 = Haemoglobin \times 1.34 \times SaO_2
\end{equation}
and
cardiac output, the primary cardiac factor that influences $\dot{V}_{O_2}$, is:
\begin{equation} \label{eq:CO=SVxHR}
	Cardiac\ Output = Stroke\ Volume \times Heart\ Rate
\end{equation}
Stroke volume is in turn influenced by ventricular function and end-diastolic volumes. 
The heart rate response to exercise is discussed in section \ref{sec:heart_rate} on page \pageref{sec:heart_rate}.

@@

The detrimental effects of jaundice on the heart and the circulatory system has been recognised for over 150 years now. 
King and co-workers reported that the intravenous injection of bile caused bradycardia, hypotension and eventually death in dogs \parencite{king_effect_1909}. 
The concept of a \textit{'jaundiced heart'} was fist put forth by Green and co-workers in 1986 \parencite{green_jaundiced_1986}. 
They performed choledocho-caval anastomoses in 5 dogs and studied cardiac function before and 2 weeks after this procedure. 
They reported that 'cholemia' was associated with impaired left ventricular function and blunted response to sympathomimmetic agents. 
Similar findings have been reported by other authors in animal studies \parencite{binah_obstructive_1985,bomzon_systemic_1986}.

@@

More recently, TNF-$\alpha$ levels have been reported to mediate cardiac dysfunction in animal studies and treatment with an anti-TNF-$\alpha$ agent restored myocardial contractility \parencite{yang_mechanisms_2010}. 
Obstructive jaundice was also associated with systemic hypotension and there is increasing evidence that some of this may be mediated by bile acid receptors on the vascular tree \parencite{green_systemic_1995, lefebvre_role_2009}. 
Bile acids can cause vasodilation by decreasing arterial tone and this may partly explain some of the haemodynamic adverse events that occur after surgery in jaundiced patients.

@@

A recent randomised controlled trial of 202 patients reported that there was no significant difference in post-operative morbidity (47\% vs, 37\%, p=0.14), mortality or length of stay between the group that underwent preoperative biliary drainage and the group that underwent early surgery \parencite{van_der_gaag_preoperative_2010}. 
There was a high incidence of procedure related complications in the drainage group (46\%). 
The high incidence of procedure related complications, the use of plastic stents, the 25\% failure to drain at first attempt and exclusion of patients with serum bilirubin $>$ 250 $\mu$mol/L were significant limitations of the study. 
However, the study does show that early surgery in the jaundiced patient with serum bilirubin $<$ 250 $\mu$mol/L is safe and not associated with increased morbidity or mortality as was once feared.

@@

The paucity of good quality evidence on the role of preoperative biliary drainage before pancreatic resection \parencite{wang_preoperative_2008} and the limitations in our understanding of the complex patho-physiological effects of jaundice on multiple organ systems has resulted in persistent wide variation in the perioperative management of jaundiced patients across different centres.

@@

The impact of obesity on the rate of postoperative complications remains controversial. 
It has been a commonly held belief that obese patients had greater incidence of complications. 
Some of the adverse outcomes such as wound complications and respiratory dysfunction where considered to be the direct result of the increased body mass index while other complications where related to the associated comorbidities such as cardiovascular disease, obstructive sleep apnea or diabetes mellitus.

@@

Tsai and co-workers reported that while obese patients had similar tumour characteristics to non-obese patients and comparable perioperative outomes, long-term survival was better in obese subjects \parencite{tsai_impact_2010}. 
This study of 795 patients undergoing potentially curative surgery for pancreatic adenocarcinoma at the Johns Hopkins Medical Institutions in the United States is in contrast to other similar studies. 
The role of obesity on long-term survival after pancreaticoduodenectomy for pancreatic cancer is less clear with more recent studies suggesting that there was no link between body mass index and survival \parencite{dandona_influence_2011, benns_impact_2009}.

>>>>

\subsection{Postoperative care}
All patients were routinely admitted to the Surgical High Dependency Unit unless intra-operative events necessitated admission to the Intensive Care Unit. 
A standardised regimen of intravenous fluids, naso-jejunal feeding, mobilisation and physiotherapy was implemented in all patients. 
Standard physiological parameters including haemodynamic measurements, renal function and arterial blood gases were used to monitor adequate end organ perfusion. 
All patients received proton pump inhibitors and Octreotide. 
Patients were discharged to the general surgical ward as early as possible. 
Amylase content in the drain fluid was measured on the third post-operative day to aid in the diagnosis of post-operative pancreatic fistula. 
Surgical drains and naso-gastric/naso-jejunal tubes were removed at the surgeon's discretion. 
Patients were discharged when medically fit.

@@

Oettle and co-workers reported on the results of a European multi-center, randomised controlled phase 3 trial comparing the results of surgery versus surgery and adjuvant chemotherapy with 6 cycles of gemcitabine in patients with pancreatic ductal adenocarcinoma \parencite{oettle_adjuvant_2007}. The median disease free survival in patient who received gemcitabine chemotherapy was 13.4 months while it was only 6.9 months in the surgery-only group. However, there was no difference in overall survival.

@@

Several studies have attempted to evaluate various surgical strategies in improving perioperative outcomes after pancreaticoduodenectomy. 
A prospective, multi-center, randomised controlled trial of 170 patients undergoing pancreaticoduodenectomy did not demonstrate any significant difference between a Whipple's classical pancreaticoduodenectomy  and pylorus preserving pancreaticoduodenectomy \parencite{tran_pylorus_2004}. 
A Cochrane review in 2008 reported on 7 randomised controlled trials comparing classical pancreaticoduodenectomy with pylorus-preserving pancreaticoduodenectomy  and found no difference in morbidity, mortality or survival \parencite{diener_pancreaticoduodenectomy_2008}.

@@

In spite of a lack of any clear evidence of efficacy, octreotide continues to be used in many centres including ours to reduce the incidence of postoperative pancreatic fistula. 
While some early randomised trials reported the benefits of octreotide in decreasing the risk of postoperative pancreatic fistula \parencite{montorsi_efficacy_1995,nakatsuka_octreotide_2000}, other more recent trials have not reproduced these results \parencite{lowy_prospective_1997, yeo_does_2000, kollmar_prophylactic_2008}.

@@

$\dot{V}_{O_2}$ or oxygen uptake is measured breath-by-breath using digital analysis of the inspired and expired gases. 
This is then averaged, usually over time, to smooth-out any significant breath-by-breath variation. $\dot{V}_{O_2}$ increases with increasing work load and is influenced by several factors that have a role in the transport and utilisation of oxygen. 
These may be broadly classified as cardiac, pulmonary, circulatory and tissue factors. 
Some of the factors are encompassed in the following formula for $\dot{V}_{O_2}$.
\begin{equation} \label{eq:oxygen_consumption}
	\dot{V}_{O_2} = CaO_2 \times Cardiac\ Output
\end{equation}
where $CaO_2$ is $O_2$ content per ml of blood and is defined by,
\begin{equation} \label{eq:CaO2}
	CaO_2 = Haemoglobin \times 1.34 \times SaO_2
\end{equation}
and
cardiac output, the primary cardiac factor that influences $\dot{V}_{O_2}$, is:
\begin{equation} \label{eq:CO=SVxHR}
	Cardiac\ Output = Stroke\ Volume \times Heart\ Rate
\end{equation}
Stroke volume is in turn influenced by ventricular function and end-diastolic volumes. 
The heart rate response to exercise is discussed in section \ref{sec:heart_rate} on page \pageref{sec:heart_rate}.

@@

The detrimental effects of jaundice on the heart and the circulatory system has been recognised for over 150 years now. 
King and co-workers reported that the intravenous injection of bile caused bradycardia, hypotension and eventually death in dogs \parencite{king_effect_1909}. 
The concept of a \textit{'jaundiced heart'} was fist put forth by Green and co-workers in 1986 \parencite{green_jaundiced_1986}. 
They performed choledocho-caval anastomoses in 5 dogs and studied cardiac function before and 2 weeks after this procedure. 
They reported that 'cholemia' was associated with impaired left ventricular function and blunted response to sympathomimetic agents. 
Similar findings have been reported by other authors in animal studies \parencite{binah_obstructive_1985,bomzon_systemic_1986}.

@@

More recently, TNF$\alpha$ levels have been reported to mediate cardiac dysfunction in animal studies and treatment with an anti-TNF$\alpha$ agent restored myocardial contractility \parencite{yang_mechanisms_2010}. 
Obstructive jaundice was also associated with systemic hypotension and there is increasing evidence that some of this may be mediated by bile acid receptors on the vascular tree \parencite{green_systemic_1995, lefebvre_role_2009}. 
Bile acids can cause vasodilation by decreasing arterial tone and this may partly explain some of the haemodynamic adverse events that occur after surgery in jaundiced patients.

@@

A recent randomised controlled trial of 202 patients reported that there was no significant difference in post-operative morbidity (47\% vs 37\%, p=0.14), mortality or length of stay between the group that underwent preoperative biliary drainage and the group that underwent early surgery \parencite{van_der_gaag_preoperative_2010}. 
There was a high incidence of procedure related complications in the drainage group (46\%). 
The high incidence of procedure related complications, the use of plastic stents, the 25\% failure to drain at first attempt and exclusion of patients with serum bilirubin $>$ 250 $\mu$mol/L were significant limitations of the study. 
However, the study does show that early surgery in the jaundiced patient with serum bilirubin $<$ 250 $\mu$mol/L is safe and not associated with increased morbidity or mortality as was once feared.

@@

The paucity of good quality evidence on the role of preoperative biliary drainage before pancreatic resection \parencite{wang_preoperative_2008} and the limitations in our understanding of the complex pathophysiological effects of jaundice on multiple organ systems has resulted in persistent wide variation in the perioperative management of jaundiced patients across different centres.

@@

The impact of obesity on the rate of postoperative complications remains controversial. 
It has been a commonly held belief that obese patients had greater incidence of complications. 
Some of the adverse outcomes such as wound complications and respiratory dysfunction where considered to be the direct result of the increased body mass index while other complications where related to the associated comorbidities such as cardiovascular disease, obstructive sleep apnoea or diabetes mellitus.

@@

Tsai and co-workers reported that while obese patients had similar tumour characteristics to non-obese patients and comparable perioperative outcomes, long-term survival was better in obese subjects \parencite{tsai_impact_2010}. 
This study of 795 patients undergoing potentially curative surgery for pancreatic adenocarcinoma at the Johns Hopkins Medical Institutions in the United States is in contrast to other similar studies. 
The role of obesity on long-term survival after pancreaticoduodenectomy for pancreatic cancer is less clear with more recent studies suggesting that there was no link between body mass index and survival \parencite{dandona_influence_2011, benns_impact_2009}.
